---
title: "Intro_R_Rmd"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Download and read in the datafile "./quant_methods/data/tgpp.csv" from the class
website. This dataset represents the vascular plant species richness that was
collected from the Tallgrass Prairie Preserve from 10 x 10 m quadrats. Species
richness is simply the number of species that occur within a quadrat. 

```{r url import}
tgpp <- read.csv('https://raw.githubusercontent.com/dmcglinn/quant_methods/gh-pages/data/tgpp.csv')
```
Read the data into R, note this datafile has a header (i.e., it has column
names) unlike the example we examined in class.  

1. What are the names of the columns in this dataset?
```{r head}
head(tgpp,1)
```
plot, year, record_id, corner, scale, richness, easting, northing, slope, ph, yrsslb

2. How many rows and columns does this data file have?
```{r dim}
dim(tgpp)
```
11 columns and 4080 rows

3. What kind of object is each data column? Hint: checkout the function sapply().
```{r sapply}
sapply(tgpp,class)
```
  Plot "integer" year "integer" record_id "integer" corner "integer" scale "numeric" richness "integer" easting "integer" northing "integer" slope "integer" ph "numeric" yrsslb "numeric"
  
4. What are the values of the the datafile for rows 1, 5, and 8 at columns 3, 
7, and 10
```{r values of rows 1,5,8 at columns 3,7,10}
tgpp[1,3]
tgpp[1,7]
tgpp[1,10]
tgpp[5,3]
tgpp[5,7]
tgpp[5,10]
tgpp[8,3]
tgpp[8,7]
tgpp[8,10]
```


                      
[1 (Row 1 columns 3,7,10)    187    727000    6.9]
[5 (Row 5 columns 3,7,10)    191    727000    6.9]
[8 (Row 8 columns 3,7,10)    194    727000    6.9]


5. Create a pdf of the relationship between the variables "scale" and "richness". 
Scale is the area in square meters of the quadrat in which richness was
recorded. Be sure to label your axes clearly, and choose a color you find
pleasing for the points. To get a list of available stock colors use the
function colors(). Also see this link: 
<http://research.stowers-institute.org/efg/R/Color/Chart/index.htm>.

```{r plot}
plot(tgpp$scale,tgpp$richness,xlab="Scale (area in square meters)",ylab="Richness",col="dodgerblue",main="Relationship between Scale and Richness")
```

6. What happens to your plot when you set the plot argument log equal to 'xy'.
`plot(..., log='xy')`
```{r log equals to "xy"}
plot(tgpp$scale,tgpp$richness,xlab="Scale",ylab="Richness",col="dodgerblue", log="xy")
```
4 y values <= 0 omitted from logarithmic plot



